---
title: "Predicting Weight from Fitness Patterns"
author: "Isobel Bodefeld, Michael Tao, Wei-jet Chiew"
format: html
---

## Introduction

We found our data on Kaggle from the author Vala Khorasani. According to Khorasani, this dataset was generated using simulated data based on "realistic gym exercise patterns." With 973 observations, this data is related to key performance indicators such as heart rate and calories burned along with information relating to demographics and experience levels.

Outcome to predict: Weight (in kilograms)

Predictors of Interest:

-   Age (in years)

-   Gender (Male or Female)

-   Height (in meters)

-   Avg_BPM (beats per minute during workout sessions)

-   Session_Duration (duration of each workout session in hours)

-   Calories_Burned (total calories burned during each session)

-   Workout_Type (Cardio, Strength, Yoga, HIIT)

    -   Cardio: 255 observations

    -   Strength: 258 observations

    -   Yoga: 239 observations

    -   HIIT: 221 observations

-   Fat_Percentage (body fat percentage)

-   Water_Intake (daily water intake in liters)

-   Workout_Frequency (number of workout sessions per week as days/week)

-   Experience_Level (beginner 1 to expert 3)

-   BMI (Body Mass Index, calculated from height and weight)

## Research Questions

Can Body Weight(kg) be determined by Workout Frequency, Workout Type, Height, Calories Burned per Workout, Fat Percentage, and Avg BPM?

Can the Calories Burned per Workout be predicted by the Session Duration of the Workout?

Can Body Fat Percentage be predicted by Experience Level, Workout Frequency, Water Intake, and Avg BPM during the workout?

## Data Exploration

```{r}
data <- read.csv("gym_members_exercise_tracking.csv")
head(data)
```

```{r}
workout_type_counts <- table(data$Workout_Type)
workout_type_counts
```

```{r}
# Visualizing Research Question 1
library(GGally)

rq_1 <- subset(data, select = c(Weight, Workout_Frequency, Workout_Type, Calories_Burned, Height, Avg_BPM, Fat_Percentage))
ggpairs(rq_1, lower = list(combo = "box_no_facet"))
```
```{r}
# Visualizing Research Question 2
rq_2 <- subset(data, select = c(Calories_Burned, Session_Duration))
ggpairs(rq_2, lower = list(combo = "box_no_facet"))
```

```{r}
# Visualizing Research Question 3
rq_3 <- subset(data, select = c(Fat_Percentage, Experience_Level, Workout_Frequency, Water_Intake, Avg_BPM))
ggpairs(rq_3, lower = list(combo = "box_no_facet"))
```
## Multiple Linear Regression Model

(Your text here)

## Improving the Model

(Your text here)

## Formal Hypothesis Tests

(Your text here)

## Robustness of Results

(Your text here)

## Conclusions

(Your text here)
